---
title: "Practicals"
author: "Lorène Jimenez"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Weekly practical 1
Lorène Jimenez, spring 2022

## Master thesis in Budongo

## Introduction
1. State of the art
2. Research question
3. Hypotheses

### 1: State of the art
Many mammals produce food calls to inform conspecifics that they found a new source of food. Chimpanzees represent a species that use this mechanism often. The exact function of this behavior is not yet known. Different hypotheses have been raised: for example, it has been suggested that food calls allow individuals to strengthen their social bonds. However, this hypothesis is not enough to explain the existence of this behavior which imply a considerable cost for the caller. An important question to ask is: how is this behavior induced in individuals if it does not imply a direct benefit for them? One possibility is that this behavior is induced by group control from other individuals because it benefits the group as a whole. More precisely, I will try to show that chimpanzees use negative direct reciprocity as a mechanism to "punish" individuals that did not produce a food call when they could have. 

### 2: Research question
Is direct negative reciprocity used to induce cooperative behavior in food-sharing contexts in chimpanzees?

### 3: Hypotheses
1. If an individual does not produce a food call when he had the occasion, and if others discover it, they produce an aggressive answer (negative direct reciprocity). 
2. The individual that undergoes this aggressive answer will increase his food calling in the near future. 

## Methods
### *Step 1*
Data collection will occur in **Budongo Conservation Field Station**. It will last 6 months, from August 2022 to February 2023. 

### *Step 2*
I will write my thesis during the second semester of my second year, spring 2023. 

![](https://media.istockphoto.com/photos/chimpanzee-sitting-in-a-tree-wildlife-shot-gombetanzania-picture-id1320018407?b=1&k=20&m=1320018407&s=170667a&w=0&h=q0Ap73ONQHJO2owq8sZE0zuW9dZHgQDGagewXDgsy08=)

Here is a table for a preloaded dataset in r: 

```{r HairEyeColor}
library(knitr)
kable(HairEyeColor)
```




# weekly practical 2
## Exercice 1: r preloaded dataset

For this exercice, I have chosen to work with the preloaded data `chickwts` from R. This dataset describes the weight of chicken according to type of food they ate: 

```{r}
library(knitr)
kable(chickwts)
```

Here is the structure of this dataset: 
```{r}
data(chickwts)
head(chickwts, 6)
str(chickwts)
```

What kind of R data type is it?

### weight
weight is a numerical variable describing the weight of chickens. 

### feed
Feed is a factor variable describing the type of food ingested by chickens. There are 6 possibilities. 

## Exercice 2: dataset from an external source
### first dataset
For this exercise I have decided to use a dataset from a [pudding article](https://pudding.cool/2019/04/vogue/) about the skin tones of models in the cover of Vogue magazine. 

```{r}
face <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/vogue/faces.csv")
head(face, 6)
str(face)
```
This dataset contains 4 variables. The first one, `date`, gives the date of the publication of the magazine. The second one, `model`, gives the name of the model on the cover. `tone` is the hex code for the skin color of each model. Finally, `l` represents the lightness of the skin tone for each model, converted to hsl format. The first variable is a date-type variable, `model` and `tone` are text variables and the last one is float variable. 

###second data set
This dataset also comes from a [pudding article](https://pudding.cool/2017/07/cetaceans/). 
```{r}
acquisitions <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/cetaceans/acquisitions.csv")
head(acquisitions, 6)
str(acquisitions)

```
This dataset contains 4 variables, all number variables in r. `AcqYear` gives the year the animal joined the community. `Born` gives the number of cetaceans that were born in captivity during a given year. `Capture` gives the number of cetaceans that were captures during a certain year. Finally, `rescue` gives the number of cetaceans that were rescued during a given year. 

All variables are quantitative discrete variables. 



# Weekly practical 3

We will be using the `gapminder` data set from the `dslabs` package. 
```{r}
update.packages("rlang")
library(dslabs)
library(dplyr)
library(knitr)
```


```{r filter}
gapminder2 <- gapminder %.%
  filter(country %in% c("Switzerland","France"), year>=1980)
```
```{r}
install.packages("rlang")
library(rlang)
```


```{r}
library(dslabs)
data(murders)
library(dplyr)
library(tidyverse)
library(knitr)
murders %>% head() %>% kable()
```

